---
title: "DATA 606 Data Project Proposal"
author: "Amanda Arce"
date: "October 22, 2018"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github

---



### Data Preparation

```{r setup, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
# load data
library(tidyverse)
library(readr)
library(psych)
library(corrgram)

```


### Research question 

**You should phrase your research question in a way that matches up with the scope of inference your dataset allows for.**

What is the deathrate based off disease, race, and age

### Cases 

**What are the cases, and how many are there?**

708 total causes - showing year, leading cause of death, ethnicity, and death/rates


### Data collection 

**Describe the method of data collection.**
Data was gathered from NYC Open Data


### Type of study 

**What type of study is this (observational/experiment)?**
Obervational 
 

### Data Source 

**If you collected the data, state self-collected. If not, provide a citation/link.**

https://data.cityofnewyork.us/Health/New-York-City-Leading-Causes-of-Death/jb7j-dtam


```{r}

leadingCauses <- read_csv("New_York_City_Leading_Causes_of_Death.csv")
head(leadingCauses, 10)
```




### Dependent Variable

**What is the response variable? Is it quantitative or qualitative?**

The dependent variable is Death rates.  The death rate is quantitative data.


### Independent Variable

**You should have two independent variables, one quantitative and one qualitative.**

The indepedent variables are disease, race, and age.


### Relevant summary statistics 

**Provide summary statistics for each the variables. Also include appropriate visualizations related to your research question (e.g. scatter plot, boxplots, etc). This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**

```{r}
leadingCauses$Deaths[leadingCauses$Deaths=="."] <- NA
leadingCauses$`Death Rate`[leadingCauses$`Death Rate`=="."] <- NA
leadingCauses$`Age Adjusted Death Rate`[leadingCauses$`Age Adjusted Death Rate`=="."] <- NA
leadingCauses <- na.omit(leadingCauses)


leadingCauses$Deaths <- as.numeric(leadingCauses$Deaths)
leadingCauses$`Death Rate` <- as.numeric(leadingCauses$`Death Rate`)
leadingCauses$`Age Adjusted Death Rate` <- as.numeric(leadingCauses$`Age Adjusted Death Rate`)
```


```{r message=FALSE, warning=FALSE}
summary(leadingCauses)
```


```{r message=FALSE, warning=FALSE, paged.print=FALSE}
describe(leadingCauses)
```

```{r}
ggplot(data = leadingCauses, aes(leadingCauses$Year, leadingCauses$Deaths)) + 
  stat_summary(fun.y = sum,
               geom = "bar", aes(fill = Year))
```

```{r}
ggplot(data = leadingCauses, aes(leadingCauses$`Race Ethnicity`, leadingCauses$Deaths)) + 
  stat_summary(fun.y = sum,
               geom = "bar", aes(fill = `Race Ethnicity`))
```

```{r}
leadingCauses %>%
  group_by(leadingCauses$`Race Ethnicity`) %>%
  summarize(mean = mean(Deaths), sd = sd(Deaths), median = median(Deaths))

```



```{r}

ggplot(data = leadingCauses, aes(leadingCauses$`Leading Cause`, leadingCauses$Deaths)) + 
  stat_summary(fun.y = sum,
               geom = "bar", aes(fill = `Leading Cause`)) +
  theme(legend.position="none")
```


```{r}
nums <- unlist(lapply(leadingCauses, is.numeric)) 
nums1 <- leadingCauses[ , nums]
cor(nums1)
```

```{r}
corrgram(nums1, order=TRUE, 
         lower.panel=panel.shade,
        upper.panel=panel.pie, 
        text.panel=panel.txt,
        main="Data")
```



















#Reference:https://data.cityofnewyork.us/Health/New-York-City-Leading-Causes-of-Death/jb7j-dtam